\bf Tearing Mode Dynamics in Tokamak Plasmas

### 2021-11-22 Documentation/TearingModeTheory.tex
Lastest version

\section{Reduced Drift-MHD Resonant  Response Model}\label{ff}
\subsection{Introduction}
Equations~(\ref{e202})--(\ref{e206}) contain two quantities---namely, the reconnection time, $\tau_s$, and the unperturbed
natural frequency, $\omega_{s\,0}$---that can only be determined by solving for the plasma response in the radially-thin resonant layer surrounding the resonant surface. Unfortunately, determining the resonant plasma response is not a straightforward task. The
problem is that the small characteristic lengthscale in the resonant region (i.e., the resonant layer width) causes the fundamental ordering parameters
$\delta_e$ and $\delta_i$ (see Section~\ref{s3.9}) to take values that are of order unity. Consequently, most of the simplifications described in
Section~\ref{lowest} are not applicable in the resonant layer. In particular, many of the terms in the Braginskii equations (see Section~\ref{sbrag}) that are negligible in the outer region (e.g., plasma resistivity) have to be taken into account in the resonant layer.

A tearing mode is a modified shear-Alfv\'{e}n wave (Hazeltine \& Meiss 1985), and has very little connection with a compressible-Alfv\'{e}n wave. For instance, the resonance condition, ${\bf k}\cdot{\bf B}=0$, which determines the position of the resonant
surface, is identical to the resonance condition for a  shear-Alfv\'{e}n wave, but differs substantially from that of a  compressible-Alfv\'{e}n wave (Fitzpatrick 2015). Hence, one very effective way of simplifying the resonant layer equations is to remove the
physics of  compressible-Alfv\'{e}n waves from them altogether, in the process converting a drift-magnetohydrodynamical (MHD) response model
into a so-called {\em reduced}\/ drift-MHD response model (Strauss 1976; Hazeltine, Kotschenreuther \& Morrison 1985; 
Furuya, Yagi \& Itoh 2003). The aim of
this section is to describe the reduction process (which essentially boils down to arguing that the divergence of the
plasma flow is small), and to derive the reduced drift-MHD equations that will subsequently be used to model the resonant
response of the plasma.

@@

Following Hazeltine \& Meiss 1992, if we neglect electron inertia and electron viscosity  in the electron equation of motion, (\ref{e3.79b}), then
we obtain the {\em generalized Ohm's law}: 
\begin{equation}\label{e220}
{\bf E} + {\bf V}\times {\bf B} + \frac{1}{e\,n_0}\!\left[\nabla p + \frac{1}{1+\tau}\left(\frac{0.71\,\eta_e\,\tau}{1+\eta_e}-1\right)({\bf b}\cdot\nabla p)\,{\bf b} -{\bf j}\times{\bf b}\right] = \eta_\parallel\,{\bf j}_\parallel + \eta_\perp\,{\bf j}_\perp.
\end{equation}
Here, $\eta_\parallel=1/\sigma_\parallel(r_s)$ and $\eta_\perp=1/\sigma_\perp(r_s)$ are the parallel and perpendicular plasma
resistivities,  respectively, at the resonant surface. (See Section~\ref{sbrag}.) We have neglected the cross term in Equation~(\ref{e3.91a})
because it is a factor $({\mit\Omega}_e\,\tau_e)^{-1}$  smaller than $\nabla p$. Incidentally, if we assume that all
flow velocities are similar in magnitude to the diamagnetic velocity then the neglect of
electron inertia and electron viscosity in the previous equation is justified because the corresponding terms are factors
$(\rho_e/r_s)^2$ and $({\mit\Omega}_e\,\tau_e)\,(\rho_e/r_s)^2$, respectively, smaller than the leading order terms
in the equation.

@@

Equations~(\ref{e220})--(\ref{e223}) are known collectively as the {\em drift-MHD equations},\footnote{The  `drift' appellation  merely indicates that we are taking diamagnetic flows into account.} and contain both shear-Alfv\'{e}n
and compressible-Alfv\'{e}n wave dynamics. The system of equations is completed by a subset of Maxwell's equations:
\begin{align}
\nabla\cdot{\bf B} &= 0,\\[0.5ex]
\nabla\times {\bf E} &= - \frac{\partial {\bf B}}{\partial t},\\[0.5ex]
\mu_0\,{\bf j} &= \nabla \times {\bf B}.\label{e226}
\end{align}

@@

It is convenient to adopt the following normalization scheme that renders all quantities in the drift-MHD equations dimensionless: 
$\hat{\nabla} = l\,\nabla$, $\hat{d}_i=d_i/l$, $\hat{t} = t/(l/V_A)$, $\hat{\bf B} = {\bf B}/B_z$, $\hat{\bf E} = {\bf E}/(B_z\,V_A)$,
$\hat{\bf j} = {\bf j}/(B_z/\mu_0\,l)$, $\hat{\eta}_{\parallel,\perp} = \eta_{\parallel,\perp}/(\mu_0\,V_A\,l)$, $\hat{\bf V} = {\bf V}/V_A$,
$\hat{{\bf V}}_{\ast,i}= {\bf V}_{\ast,i}/V_A$, $\hat{V}_{\parallel\,i}= V_{\parallel\,i}/V_A$, $\hat{p} = p/(B_z^{\,2}/\mu_0)$, $\hat{p}_0 = p_0/(B_z^{\,2}/\mu_0)$, $\hat{\bpi}_{i} = \bpi_{i}/(B_z^{\,2}/\mu_0)$,
$\hat{\bf q} = {\bf q} /(B_z^{\,2}\,V_A/\mu_0)$. Equations~(\ref{e220})--(\ref{e226}) yield
\begin{align}\label{e227}
\hat{\bf E} + \hat{\bf V}\times \hat{\bf B} + \hat{d}_i\!\left[\hat{\nabla} \hat{p} + \frac{1}{1+\tau}\left(\frac{0.71\,\eta_e\,\tau}{1+\eta_e}-1\right)({\bf b}\cdot\hat{\nabla} \hat{p})\,{\bf b} + \hat{\nabla}\cdot\hat{\bpi}_e-\hat{\bf j}\times\hat{\bf b}\right] &= \hat{\eta}_\parallel\,\hat{\bf j}_\parallel+ \hat{\eta}_\perp\,\hat{\bf j}_\perp,\\[0.5ex]
\frac{\partial \hat{\bf V}}{\partial \hat{t}} 
+ (\hat{\bf V}\cdot\hat{\nabla} )\hat{\bf V} + 
\frac{1}{1+\tau}\,(\hat{\bf V}_\ast\cdot\hat{\nabla})\hat{\bf V}_E
+ \hat{\nabla} \hat{p} + \hat{\nabla}\cdot\hat{\bpi}_i -
 \hat{\bf j}\times \hat{\bf B} &= 0,
 \\[0.5ex]
\frac{\partial \hat{p}}{\partial \hat{t}} + \hat{\bf V}\cdot\hat{\nabla} \hat{p} + {\mit\Gamma}\,\hat{p}\,\hat{\nabla}\cdot\hat{\bf V} + \frac{2}{3}\,\hat{\nabla}\cdot \hat{\bf q}&= 0,\label{e229}
\end{align}
and
\begin{align}
\hat{\nabla}\cdot\hat{\bf B} &= 0,\label{e230}\\[0.5ex]
\hat{\nabla}\times \hat{\bf E} &= - \frac{\partial \hat{\bf B}}{\partial \hat{t}},\label{e231}\\[0.5ex]
\hat{\bf j} &= \hat{\nabla} \times \hat{\bf B}.\label{e232}
\end{align}
Finally,
\begin{align}
\hat{\bf V}_E &= \hat{\bf E}\times {\bf b},\\[0.5ex]
\hat{\bf V} &= \hat{\bf V}_E + \hat{V}_{\parallel\,i}\,{\bf b},\\[0.5ex]
\hat{\bf V}_\ast &= \hat{d}_i\,{\bf b}\times \hat{\nabla} \hat{p},\\[0.5ex]
\hat{\bf V}_i &= \hat{\bf V} + \frac{\hat{\bf V}_{\ast\,i}}{1+\tau}.
\end{align}

@@

\subsection{Discussion}
Equations~(\ref{e264})--(\ref{e269}) constitute our reduced, four-field, drift-MHD, resonant, plasma response model. Our model is very similar
to the original four-field model derived by Hazeltine, et al.\ (Hazeltine, et al.\ 1985). Equation~(\ref{e264}) is the {\em generalized
Ohm's law}\/ that governs the time evolution of the (normalized) helical magnetic flux, $\hat{\psi}$. Equation~(\ref{e265}) is the {\em parallel
ion vorticity equation}\/ that governs the  time evolution of the (normalized) parallel ion vorticity, $U$. 
Equation~(\ref{e266}) is the {\em electron number density continuity equation}\/ that 
governs the time evolution of the (normalized) electron number density, $N$. Finally, Equation~(\ref{e267}) is the
{\em ion parallel equation of motion}\/ that governs the time evolution of the  (normalized) parallel ion velocity, $V$.

@@

\section{Linear Resonant  Response Model}\label{linear}
\subsection{Introduction}
The aim of this section is to employ the reduced, drift-MHD, four-field, resonant, plasma response model derived in Section~\ref{ff} to determine the {\em linear}\/ response of the resonant layer to external perturbations.

@@

\subsection{Derivation of Linear Layer Equations}\label{s5.3}
Suppose that the system is perturbed by a tearing mode with $m$ periods in the poloidal direction, and
$n$ periods in the toroidal direction. In accordance with Equations~(\ref{e287})--(\ref{e290}), and (\ref{e292})--(\ref{e293}), we can write
\begin{align}\label{e295}
\hat{\psi}(\hat{x},\zeta,t) &= \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ \tilde{\psi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
\phi(\hat{x},\zeta,t) &=-\hat{V}_E\,\hat{x}+ \tilde{\phi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
N(\hat{x},\zeta,t) &= -\hat{V}_\ast\,\hat{x}+\left(\frac{1+\tau}{\tau}\right)\!\tilde{N}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
V(\hat{x},\zeta,t) &= \hat{V}_\parallel + \left(\frac{1+\tau}{\tau}\right)\!\tilde{V}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
J(\hat{x},\zeta,t) &=-\left(\frac{2}{s_s}-1\right)\!\frac{1}{\hat{L}_s}+ \hat{\nabla}^2\tilde{\psi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
U(\hat{x},\zeta,t) &=  \hat{\nabla}^2\tilde{\phi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\label{e300}
\end{align}
where $\hat{t}=V_A\,t/r_s$, $\hat{\omega}=r_s\,\omega/V_A$, and $\omega$ is the frequency of the tearing mode in the laboratory frame. 
Substituting Equations~(\ref{e294})--(\ref{e300}) into the four-field model, (\ref{e264})--(\ref{e269}), and only retaining terms that
are first order in perturbed quantities, we obtain the following set of linear layer equations:
\begin{align}\label{e301}
-{\rm i}\,[\omega-\omega_E -(1+\lambda_e)\,\omega_{\ast\,e}]\,\tau_H\,\tilde{\psi} &= -{\rm i} \,\hat{x}\,[\tilde{\phi}-(1+\lambda_e)\,\tilde{N}] + S^{-1}\,\hat{\nabla}^2\tilde{\psi},\\[0.5ex]
-{\rm i}\,(\omega-\omega_E-\omega_{\ast\,i})\,\tau_H\,\hat{\nabla}^2\tilde{\phi} &= -{\rm i}\,\hat{x}\,\hat{\nabla}^2\tilde{\psi}+S^{-1}\,P_\varphi\,\hat{\nabla}^4\!\left(\tilde{\phi} + \frac{\tilde{N}}{\tau}\right),\\[0.5ex]
-{\rm i}\,(\omega-\omega_E)\,\tau_H\,\tilde{N} &= -{\rm i}\,\omega_{\ast\,e}\,\tau_H\,\tilde{\phi}-{\rm i}\,c_\beta^{\,2}\,\hat{x}\,\tilde{V} - {\rm i}\!\left(\frac{\tau}{1+\tau}\right)\hat{d}_\beta^{\,2}\,\hat{x}\,\hat{\nabla}^2\tilde{\psi} \\[0.5ex]
&\phantom{=}
+S^{-1}\,P_\parallel\!\left(m\,\omega_{\ast\,e}\,\tau_H\,\hat{x}\,\tilde{\psi}-\frac{m}{\hat{L}_s}\,\hat{x}^2\,\tilde{N}\right)
+ S^{-1}\,P_\perp\,\hat{\nabla}_\perp^2 \tilde{N},\nonumber\\[0.5ex]
-{\rm i}\,(\omega-\omega_E)\,\tau_H\,\tilde{V} &= {\rm i}\,\omega_{\ast\,e}\,\tau_H\,\tilde{\psi} - {\rm i}\,\hat{x}\,\tilde{N}
+ S^{-1}\,P_\varphi\,\hat{\nabla}^2\tilde{V}.\label{e304}
\end{align}
Here, 
\begin{equation}
\tau_H = \frac{L_s}{m\,V_A}
\end{equation}
is the {\em hydromagnetic time}, 
\begin{equation}
\omega_E = \left(\frac{m}{r_s}\right)V_E(r_s)
\end{equation}
  the 
{\em E-cross-B frequency}, 
\begin{equation}
\omega_{\ast\,e} = -\left(\frac{\tau}{1+\tau}\right)\left(\frac{m}{r_s}\right)V_\ast(r_s)
\end{equation}
  the {\em electron diamagnetic
frequency},
\begin{align}
  \omega_{\ast\,i} &= \left(\frac{1}{1+\tau}\right)\left(\frac{m}{r_s}\right)V_\ast(r_s)
  \end{align}
    the {\em ion diamagnetic
frequency}, 
\begin{equation}
S=\frac{\tau_R}{\tau_H}
\end{equation}
 the {\em Lundquist number}, 
\begin{equation}
\tau_R = \frac{\mu_0\,r_s^{\,2}}{\eta_\parallel}
\end{equation}
 the
{\em resistive diffusion time}, 
\begin{equation}
\tau_\varphi
= \frac{r_s^{\,2}}{\chi_\varphi}=\frac{r_s^{\,2}\,n_0\,m_i}{\mu_\perp(r_s)}
\end{equation} 
the {\em toroidal momentum confinement time}, and 
\begin{equation}
\tau_\perp = \frac{r_s^{\,2}}{D_\perp}
\end{equation}
 the {\em particle/energy confinement time}. Furthermore, $P_\varphi = \tau_R/\tau_\varphi$, $P_\perp = \tau_R/\tau_\perp$, 
$P_\parallel=\tau_R/\tau_\parallel$, and $\tau_\parallel =r_s^{\,2}/D_\parallel$.

@@

Suppose that the resonant layer lies in a constant-$\psi$ response regime. According to Equation~(\ref{e295}),
the perturbed magnetic flux in the layer takes the form:
\begin{equation}
\hat{\psi}(\hat{x},\zeta,t) = \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ \tilde{\psi}\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},
\end{equation}
where $\tilde{\psi}$ is a complex constant. Of course, the physical flux is the real part of the previous expression; that is,
\begin{equation}\label{e352a}
\hat{\psi}(\hat{x},\xi)= \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ |\tilde{\psi}|\,\cos\xi,
\end{equation}
where
\begin{equation}
\xi = \zeta +{\rm arg}(\tilde{\psi})-\hat{\omega}\,\hat{t} = m\,\theta-n\,\varphi +{\rm arg}(\tilde{\psi}) -\omega\,t.
\end{equation}
Now, it is clear from Equation~(\ref{e279}) that ${\bf B}\cdot\hat{\nabla}\hat{\psi}=0$. In other words, the contours of
$\hat{\psi}(\hat{x},\xi)$ map out the perturbed magnetic flux-surfaces in the vicinity of the resonant layer. 
Let $\hat{W} = 4\,(|\tilde{\psi}|\,\hat{L}_s)^{1/2}$. Equation~(\ref{e352a}) can be written
\begin{equation}
\frac{\hat{\psi}(\hat{x},\xi)}{|\tilde{\psi}|} = 8\left(\frac{\hat{x}}{\hat{W}}\right)^2 + \cos\xi.
\end{equation}
Figure~\ref{f2a} shows the contours of $\psi(\hat{x},\xi)$ specified by the previous equation. It can be seen that the
tearing mode has reconnected magnetic field-lines to produce a {\em helical magnetic island chain}\/ centered on the resonant
surface ($\hat{x}=0$). The chain has $m$ periods in the poloidal direction, $n$ periods in the toroidal direction, and propagates at the
helical phase velocity $\omega$. The maximum full (as opposed to half) radial width of the magnetic separatrix is 
\begin{equation}\label{e355s}
W =  \hat{W}\,r_s= 4\,(|\tilde{\psi}|\,\hat{L}_s)^{1/2}\,r_s = 4\left(\frac{q_s}{s_s}\,|\hat{\mit\Psi}_s|\right)^{1/2}R_0,
\end{equation}
where ${\mit\Psi}_s = r_s\,B_z\,\tilde{\psi}$ is the reconnected magnetic flux defined in Equation~(\ref{e99}), and
$\hat{\mit\Psi}_s= {\mit\Psi}_s/(R_0\,B_z)$.

@@

\item Furuya, A., Yagi, M., and Itoh, S.-I.\ 2003. {\em Linear Analysis of Neoclassical Tearing Mode

>>>>

\section{Reduced Resonant  Response Model}\label{ff}
\subsection{Introduction}
Equations~(\ref{e202})--(\ref{e206}) contain two quantities---namely, the reconnection time, $\tau_s$, and the unperturbed
natural frequency, $\omega_{s\,0}$---that can only be determined by solving for the plasma response in the radially-thin resonant layer surrounding the resonant surface. Unfortunately, determining the resonant plasma response is not a straightforward task. The
problem is that the small characteristic lengthscale in the resonant region (i.e., the resonant layer width) causes the fundamental ordering parameters
$\delta_e$ and $\delta_i$ (see Section~\ref{s3.9}) to take values that are of order unity. Consequently, most of the simplifications described in
Section~\ref{lowest} are not applicable in the resonant layer. In particular, many of the terms in the Braginskii equations (see Section~\ref{sbrag}) that are negligible in the outer region (e.g., plasma resistivity) have to be taken into account in the resonant layer.

A tearing mode is a modified shear-Alfv\'{e}n wave (Hazeltine \& Meiss 1985), and has very little connection with a compressible-Alfv\'{e}n wave. For instance, the resonance condition, ${\bf k}\cdot{\bf B}=0$, which determines the position of the resonant
surface, is identical to the resonance condition for a  shear-Alfv\'{e}n wave, but differs substantially from that of a  compressible-Alfv\'{e}n wave (Fitzpatrick 2015). Hence, one very effective way of simplifying the resonant layer equations is to remove the
physics of  compressible-Alfv\'{e}n waves from them altogether, in the process converting a 
drift-magnetohydrodynamical\footnote{The  `drift' appellation  merely indicates that we are taking diamagnetic flows into account. However, this necessarily entails treating the ion and electron flow separately. In other words, a
drift-MHD model is also a `two-fluid' model. }(MHD) response model  
into a so-called {\em reduced}\/ drift-MHD response model (Strauss 1976; Hazeltine, Kotschenreuther \& Morrison 1985). The aim of
this section is to describe the reduction process (which essentially boils down to arguing that the divergence of the
plasma flow is small), and to derive the reduced drift-MHD equations that will subsequently be used to model the resonant
response of the plasma.

@@

Following Hazeltine \& Meiss 1992, if we neglect electron inertia and electron viscosity  in the electron equation of motion, (\ref{e3.79b}), then
we obtain the {\em generalized Ohm's law}: 
\begin{equation}\label{e220}
{\bf E} + {\bf V}\times {\bf B} + \frac{1}{e\,n_0}\!\left[\nabla p + \frac{1}{1+\tau}\left(\frac{0.71\,\eta_e\,\tau}{1+\eta_e}-1\right)({\bf b}\cdot\nabla p)\,{\bf b} -{\bf j}\times{\bf B}\right] = \eta_\parallel\,{\bf j}_\parallel + \eta_\perp\,{\bf j}_\perp.
\end{equation}
Here, $\eta_\parallel=1/\sigma_\parallel(r_s)$ and $\eta_\perp=1/\sigma_\perp(r_s)$ are the parallel and perpendicular plasma
resistivities,  respectively, at the resonant surface. (See Section~\ref{sbrag}.) We have neglected the cross term in Equation~(\ref{e3.91a})
because it is a factor $({\mit\Omega}_e\,\tau_e)^{-1}$  smaller than $\nabla p$. Incidentally, if we assume that all
flow velocities are similar in magnitude to the diamagnetic velocity then the neglect of
electron inertia and electron viscosity in the previous equation is justified because the corresponding terms are factors
$(\rho_e/r_s)^2$ and $({\mit\Omega}_e\,\tau_e)\,(\rho_e/r_s)^2$, respectively, smaller than the leading order terms
in the equation.

@@

Equations~(\ref{e220})--(\ref{e223}) are known collectively as the {\em drift-MHD equations}, and contain both shear-Alfv\'{e}n
and compressible-Alfv\'{e}n wave dynamics. The system of equations is completed by a subset of Maxwell's equations:
\begin{align}
\nabla\cdot{\bf B} &= 0,\\[0.5ex]
\nabla\times {\bf E} &= - \frac{\partial {\bf B}}{\partial t},\\[0.5ex]
\mu_0\,{\bf j} &= \nabla \times {\bf B}.\label{e226}
\end{align}

@@

It is convenient to adopt the following normalization scheme that renders all quantities in the drift-MHD equations dimensionless: 
$\hat{\nabla} = l\,\nabla$, $\hat{d}_i=d_i/l$, $\hat{t} = t/(l/V_A)$, $\hat{\bf B} = {\bf B}/B_z$, $\hat{\bf E} = {\bf E}/(B_z\,V_A)$,
$\hat{\bf j} = {\bf j}/(B_z/\mu_0\,l)$, $\hat{\eta}_{\parallel,\perp} = \eta_{\parallel,\perp}/(\mu_0\,V_A\,l)$, $\hat{\bf V} = {\bf V}/V_A$,
$\hat{{\bf V}}_{\ast,i}= {\bf V}_{\ast,i}/V_A$, $\hat{V}_{\parallel\,i}= V_{\parallel\,i}/V_A$, $\hat{p} = p/(B_z^{\,2}/\mu_0)$, $\hat{p}_0 = p_0/(B_z^{\,2}/\mu_0)$, $\hat{\bpi}_{i} = \bpi_{i}/(B_z^{\,2}/\mu_0)$,
$\hat{\bf q} = {\bf q} /(B_z^{\,2}\,V_A/\mu_0)$. Equations~(\ref{e220})--(\ref{e226}) yield
\begin{align}\label{e227}
\hat{\bf E} + \hat{\bf V}\times \hat{\bf B} + \hat{d}_i\!\left[\hat{\nabla} \hat{p} + \frac{1}{1+\tau}\left(\frac{0.71\,\eta_e\,\tau}{1+\eta_e}-1\right)({\bf b}\cdot\hat{\nabla} \hat{p})\,{\bf b} + \hat{\nabla}\cdot\hat{\bpi}_e-\hat{\bf j}\times\hat{\bf B}\right] &= \hat{\eta}_\parallel\,\hat{\bf j}_\parallel+ \hat{\eta}_\perp\,\hat{\bf j}_\perp,\\[0.5ex]
\frac{\partial \hat{\bf V}}{\partial \hat{t}} 
+ (\hat{\bf V}\cdot\hat{\nabla} )\hat{\bf V} + 
\frac{1}{1+\tau}\,(\hat{\bf V}_\ast\cdot\hat{\nabla})\hat{\bf V}_E
+ \hat{\nabla} \hat{p} + \hat{\nabla}\cdot\hat{\bpi}_i -
 \hat{\bf j}\times \hat{\bf B} &= 0,
 \\[0.5ex]
\frac{\partial \hat{p}}{\partial \hat{t}} + \hat{\bf V}\cdot\hat{\nabla} \hat{p} + {\mit\Gamma}\,\hat{p}\,\hat{\nabla}\cdot\hat{\bf V} + \frac{2}{3}\,\hat{\nabla}\cdot \hat{\bf q}&= 0,\label{e229}
\end{align}
and
\begin{align}
\hat{\nabla}\cdot\hat{\bf B} &= 0,\label{e230}\\[0.5ex]
\hat{\nabla}\times \hat{\bf E} &= - \frac{\partial \hat{\bf B}}{\partial \hat{t}},\label{e231}\\[0.5ex]
\hat{\bf j} &= \hat{\nabla} \times \hat{\bf B}.\label{e232}
\end{align}
Finally,
\begin{align}
\hat{\bf V}_E &= \hat{\bf E}\times {\bf b},\\[0.5ex]
\hat{\bf V} &= \hat{\bf V}_E + \hat{V}_{\parallel\,i}\,{\bf b},\\[0.5ex]
\hat{\bf V}_\ast &= \hat{d}_i\,{\bf b}\times \hat{\nabla} \hat{p},\\[0.5ex]
\hat{\bf V}_i &= \hat{\bf V} + \frac{\hat{\bf V}_{\ast}}{1+\tau}.
\end{align}

@@

\subsection{Discussion}
Equations~(\ref{e264})--(\ref{e269}) constitute our reduced drift-MHD resonant plasma response model. Our model is very similar
to the original four-field model derived by Hazeltine, et al.\ (Hazeltine, Kotschenreuther \& Morrison 1985). Equation~(\ref{e264}) is the {\em generalized
Ohm's law}\/ that governs the time evolution of the (normalized) helical magnetic flux, $\hat{\psi}$. Equation~(\ref{e265}) is the {\em parallel
ion vorticity equation}\/ that governs the  time evolution of the (normalized) parallel ion vorticity, $U$. 
Equation~(\ref{e266}) is the {\em electron number density continuity equation}\/ that 
governs the time evolution of the (normalized) electron number density, $N$. Finally, Equation~(\ref{e267}) is the
{\em ion parallel equation of motion}\/ that governs the time evolution of the  (normalized) parallel ion velocity, $V$.

@@

\section{Linear Resonant  Response Model}\label{linear}
\subsection{Introduction}
The aim of this section is to employ the reduced resonant plasma response model derived in Section~\ref{ff} to determine the {\em linear}\/ response of the resonant layer to external perturbations.

@@

\subsection{Derivation of Linear Layer Equations}\label{s5.3}
Suppose that the system is perturbed by a tearing mode with $m$ periods in the poloidal direction, and
$n$ periods in the toroidal direction. In accordance with Equations~(\ref{e287})--(\ref{e290}), and (\ref{e292})--(\ref{e293}), we can write
\begin{align}\label{e295}
\hat{\psi}(\hat{x},\zeta,\hat{t}) &= \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ \tilde{\psi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
\phi(\hat{x},\zeta,\hat{t}) &=-\hat{V}_E\,\hat{x}+ \tilde{\phi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
N(\hat{x},\zeta,\hat{t}) &= -\hat{V}_\ast\,\hat{x}+\left(\frac{1+\tau}{\tau}\right)\!\tilde{N}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
V(\hat{x},\zeta,\hat{t}) &= \hat{V}_\parallel + \left(\frac{1+\tau}{\tau}\right)\!\tilde{V}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
J(\hat{x},\zeta,\hat{t}) &=-\left(\frac{2}{s_s}-1\right)\!\frac{1}{\hat{L}_s}+ \hat{\nabla}^2\tilde{\psi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\\[0.5ex]
U(\hat{x},\zeta,\hat{t}) &=  \hat{\nabla}^2\tilde{\phi}(\hat{x})\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},\label{e300}
\end{align}
where $\hat{t}=V_A\,t/r_s$, $\hat{\omega}=r_s\,\omega/V_A$, and $\omega$ is the frequency of the tearing mode in the laboratory frame. 
Substituting Equations~(\ref{e294})--(\ref{e300}) into the four-field model, (\ref{e264})--(\ref{e269}), and only retaining terms that
are first order in perturbed quantities, we obtain the following set of linear layer equations:
\begin{align}\label{e301}
-{\rm i}\,[\omega-\omega_E -(1+\lambda_e)\,\omega_{\ast\,e}]\,\tau_H\,\tilde{\psi} &= -{\rm i} \,\hat{x}\,[\tilde{\phi}-(1+\lambda_e)\,\tilde{N}] + S^{-1}\,\hat{\nabla}^2\tilde{\psi},\\[0.5ex]
-{\rm i}\,(\omega-\omega_E-\omega_{\ast\,i})\,\tau_H\,\hat{\nabla}^2\tilde{\phi} &= -{\rm i}\,\hat{x}\,\hat{\nabla}^2\tilde{\psi}+S^{-1}\,P_\varphi\,\hat{\nabla}^4\!\left(\tilde{\phi} + \frac{\tilde{N}}{\tau}\right),\\[0.5ex]
-{\rm i}\,(\omega-\omega_E)\,\tau_H\,\tilde{N} &= -{\rm i}\,\omega_{\ast\,e}\,\tau_H\,\tilde{\phi}-{\rm i}\,c_\beta^{\,2}\,\hat{x}\,\tilde{V} - {\rm i}\!\left(\frac{\tau}{1+\tau}\right)\hat{d}_\beta^{\,2}\,\hat{x}\,\hat{\nabla}^2\tilde{\psi} \\[0.5ex]
&\phantom{=}
+S^{-1}\,P_\parallel\!\left(m\,\omega_{\ast\,e}\,\tau_H\,\hat{x}\,\tilde{\psi}-\frac{m}{\hat{L}_s}\,\hat{x}^2\,\tilde{N}\right)
+ S^{-1}\,P_\perp\,\hat{\nabla}_\perp^2 \tilde{N},\nonumber\\[0.5ex]
-{\rm i}\,(\omega-\omega_E)\,\tau_H\,\tilde{V} &= {\rm i}\,\omega_{\ast\,e}\,\tau_H\,\tilde{\psi} - {\rm i}\,\hat{x}\,\tilde{N}
+ S^{-1}\,P_\varphi\,\hat{\nabla}^2\tilde{V}.\label{e304}
\end{align}
Here, 
\begin{equation}
\tau_H = \frac{L_s}{m\,V_A}
\end{equation}
is the {\em hydromagnetic time}, 
\begin{equation}
\omega_E = \left(\frac{m}{r_s}\right)V_E(r_s)
\end{equation}
  the 
{\em E-cross-B frequency}, 
\begin{equation}
\omega_{\ast\,e} = -\left(\frac{\tau}{1+\tau}\right)\left(\frac{m}{r_s}\right)V_\ast(r_s)
\end{equation}
  the {\em electron diamagnetic
frequency},
\begin{align}
  \omega_{\ast\,i} &= \left(\frac{1}{1+\tau}\right)\left(\frac{m}{r_s}\right)V_\ast(r_s)
  \end{align}
    the {\em ion diamagnetic
frequency}, 
\begin{equation}
S=\frac{\tau_R}{\tau_H}
\end{equation}
 the {\em Lundquist number}, 
\begin{equation}
\tau_R = \frac{\mu_0\,r_s^{\,2}}{\eta_\parallel}
\end{equation}
 the
{\em resistive diffusion time}, 
\begin{equation}
\tau_\varphi
= \frac{r_s^{\,2}}{\chi_\varphi}=\frac{r_s^{\,2}\,n_0\,m_i}{\mu_\perp(r_s)}
\end{equation} 
the {\em toroidal momentum confinement time}, and 
\begin{equation}
\tau_\perp = \frac{r_s^{\,2}}{D_\perp}
\end{equation}
 the {\em particle/energy confinement time}. Furthermore, $P_\varphi = \tau_R/\tau_\varphi$, $P_\perp = \tau_R/\tau_\perp$, 
$P_\parallel=\tau_R/\tau_\parallel$, and $\tau_\parallel =r_s^{\,2}/D_\parallel$.

@@

Suppose that the resonant layer lies in a constant-$\psi$ response regime. According to Equation~(\ref{e295}),
the perturbed magnetic flux in the layer takes the form:
\begin{equation}
\hat{\psi}(\hat{x},\zeta,\hat{t}) = \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ \tilde{\psi}\,{\rm e}^{\,{\rm i}\,(\zeta-\hat{\omega}\,\hat{t})},
\end{equation}
where $\tilde{\psi}$ is a complex constant. Of course, the physical flux is the real part of the previous expression; that is,
\begin{equation}\label{e352a}
\hat{\psi}(\hat{x},\xi)= \frac{\hat{x}^{\,2} }{2\,\hat{L}_s}+ |\tilde{\psi}|\,\cos\xi,
\end{equation}
where
\begin{equation}
\xi = \zeta +{\rm arg}(\tilde{\psi})-\hat{\omega}\,\hat{t} = m\,\theta-n\,\varphi +{\rm arg}(\tilde{\psi}) -\omega\,t.
\end{equation}
Now, it is clear from Equation~(\ref{e279}) that ${\bf B}\cdot\hat{\nabla}\hat{\psi}=0$. In other words, the contours of
$\hat{\psi}(\hat{x},\xi)$ map out the perturbed magnetic flux-surfaces in the vicinity of the resonant layer. 
Let $\hat{W} = 4\,(\hat{L}_s|\,\tilde{\psi}|)^{1/2}$. Equation~(\ref{e352a}) can be written
\begin{equation}
\frac{\hat{\psi}(\hat{x},\xi)}{|\tilde{\psi}|} = 8\left(\frac{\hat{x}}{\hat{W}}\right)^2 + \cos\xi.
\end{equation}
Figure~\ref{f2a} shows the contours of $\psi(\hat{x},\xi)$ specified by the previous equation. It can be seen that the
tearing mode has reconnected magnetic field-lines to produce a {\em helical magnetic island chain}\/ centered on the resonant
surface ($\hat{x}=0$). The chain has $m$ periods in the poloidal direction, $n$ periods in the toroidal direction, and propagates at the
helical phase velocity $\omega$. The maximum full (as opposed to half) radial width of the magnetic separatrix is 
\begin{equation}\label{e355s}
W =  \hat{W}\,r_s= 4\,(\hat{L}_s\,|\tilde{\psi}|)^{1/2}\,r_s = 4\left(L_s\,R_0\,|\hat{\mit\Psi}_s|\right)^{1/2},
\end{equation}
where ${\mit\Psi}_s = r_s\,B_z\,\tilde{\psi}$ is the reconnected magnetic flux defined in Equation~(\ref{e99}), and
$\hat{\mit\Psi}_s= {\mit\Psi}_s/(R_0\,B_z)$.

@@

%\item Furuya, A., Yagi, M., and Itoh, S.-I.\ 2003. {\em Linear Analysis of Neoclassical Tearing Mode
